\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}
\label{ch:acknowledgments}


It has been a great pleasure to be a part of the Center for Soft Matter Research at the Department of Physics at New York University. I would like to take this opportunity to express my gratitude to everyone who has assisted me during my PhD time. This thesis would not have been possible to be completed without their help.

First and foremost, I would like to acknowledge Professor David G. Grier, my thesis advisor. David has been an incredible and patient mentor throughout my graduate career. He has taught me, both consciously and unconsciously, how good experimental physics is done. David's creativity and deep insight arising from his open mind to nature and human  was contagious and motivational for me, even during tough times in the Ph.D. pursuit. His encouragement and unfailing optimism have driven me to move on steadfastly. The kindness, energy, enthusiasm, and humor he has ushered have filled the lab with warm and inspirational atmosphere all the time.  Thank you especially for the painstaking editing of this thesis! I am grateful for the great work environment that you created with your positive attitude.

I am grateful to my thesis committee, Paul Chaikin, Andrew Wray, Marc Gershaw , who supported me through the thesis process. Thank you for your questions, motivation and feedback. Special thanks to Alexander Grosberg who taught me soft matter physics.

Grierlings (members of the Grier group) are some of the most intellectual and inspiring people I have met, who will be difficult to find elsewhere. I am thankful for the overlap in my tenure I have shared with Henrique, Daniel, Chen, Mark, Aaron, Jordan, Mohammed, and Lauren. Thanks to Henrique Moyses for kick-starting my research career and teaching me all the experimental techniques as a newcomer. Thanks to Chen Wang for his help right from the beginning of my Ph.D. till today with pretty much everything including precious suggestions on my research, and job hunting. Thanks to Mark Hannel for being a great academic example everyday and coaching me all the time. I appreciate all those discussions about statistical physics, digital holography, machine learning and countless banter. Thank you for your honest feedback which helped me grow. I am grateful for the collaboration with Aaron Yevick. Without you and David Ruffner the research in this thesis would not have started.

I would like to thank everyone else in the Physics Department, especially at CSMR who made all of the paperwork look so easy. Thank you Jessica Ong and Evette Ma for your assistance. Thanks to Mark Ofitserov for making things work in the lab. Thanks also to Andrew Hollingsworth for answering all of my questions about the lab and even about basic chemistry.

I must thank my classmates and office mates who had my back throughout our time here. My time at NYU was made enjoyable in large part due to the many friends and groups that became a part of my life. I am grateful for time spent with roommates and friends, for my backpacking buddies and our memorable trips into the mountains, and oceans. Special thanks to Yishuai Xu for all those playoffs. Maybe one day we will master our backhand in tennis and also serve like Federer at court. Thanks to Daniel, Mark, Melissa, Lauren, Jordan for being the best neighbors everyday.

And finally, I would also like to say a heartfelt thank you to my parents for always believing in me and encouraging me to follow my dreams. Thanks also to Bratati whose support during the entirety of this Ph.D. is so appreciated. This dissertation would not have been possible without their warm love, continued patience, and endless support. Thank you.

